\label{sec:impacts}

The commercial success of GPU manufacturers coinciding with the boom of
AI and machine learning shows the importance of general-purpose computing on
GPUs.
%
At the same time, the public has become more aware of security and privacy
issues.
%
At present, the GPU-based systems underlying many applications, including
LLMs and other machine learning applications that process sensitive consumer
data, are quite simply ``black boxes'': the semantics of these systems are
not well-understood enough for even the most well-intentioned software
developers to provide formal guarantees.
%
Furthermore, formal {\em security} guarantees in the world of GPU programming
is essentially unexplored.
%
This proposal seeks to drastically improve the state of the art in formal
understanding of the security issues associated with GPUs, and to provide
the first formal security guarantees on these systems.
%
This will provide a benefit to society in the form of improved software
and increased confidence in the trustworthiness of AI and other cloud
applications.

This proposal will also provide numerous opportunities for student research,
including by undergraduate students.
\stefan{Update with budget numbers.}
The research will be performed at Illinois Institute of Technology, an R2
research university located in Chicago.
%
Illinois Tech undergraduates are drawn heavily from local high schools, and
the student body is quite demographically and socioeconomically diverse.
%
Both PIs have strong track records of mentoring students, including
undergraduates and students from underrepresented groups in computer science.
\textbf{PI Derakhshan} \stefan{fill in}.
%
\textbf{PI Muller} has advised~7 undergraduate research students since
joining Illinois Tech in 2020, including some who have substantially
contributed to manuscripts submitted to and published at top-tier
programming languages conferences.
%
Women and underrepresented minority students make up over one-third of his
current and former student research advisees.
